---
title: "Untitled"
author: "Yatong Li"
date: "February 8, 2015"
output: html_document
---

## Load the packages needed

```{r}
suppressMessages(library(pander))
suppressMessages(library(GEOmetadb))
suppressMessages(library(data.table))
```

## Download the GEO database

```{r}
## This will download the entire database, so can be slow
if(!file.exists("GEOmetadb.sqlite"))
{
  # Download database only if it's not done already
  getSQLiteFile()
}
```

## Problem 1

```{r}
# open connection
geo_con <- dbConnect(SQLite(), 'GEOmetadb.sqlite')

# SQL query to show title, GSE accession number, GPL accession number, manufacturer, and description
# select all expression data for HCV, Illumina platform, submitted from someone at Yale
res <- dbGetQuery(geo_con, "SELECT 
            gse.title, 
            gse.gse, 
            gpl.gpl, 
            gpl.manufacturer, 
            gpl.title 
        FROM (gse JOIN gse_gpl ON gse.gse=gse_gpl.gse) j 
        JOIN gpl ON j.gpl=gpl.gpl 
        WHERE gse.title like '%HCV%'
            AND gpl.manufacturer like '%Illumina%' 
            AND gse.contact like '%Yale%';")
res
```

Also, we know that HCV stands for hepatitis C virus so we might also search for titles that include either 'HCV' or 'hepatitis C virus'. Also, if we want even more details about the platform used, we can use the following query. When we do this we see that there is actually one more dataset whose title includes the phrase "Hepatitis C virus" rather than "HCV", so we were unable to pick it up using our first query. The additional platform description is quite detailed.

```{r}
# SQL query to show title, GSE accession number, GPL accession number, manufacturer, and description
# select all data for HCV or hepatitis C virus, Illumina platform, submitted from someone at Yale
res_v2 <- dbGetQuery(geo_con, "SELECT 
            gse.title, 
            gse.gse, 
            gpl.gpl, 
            gpl.manufacturer, 
            gpl.title,
            gpl.description
        FROM (gse JOIN gse_gpl ON gse.gse=gse_gpl.gse) j 
        JOIN gpl ON j.gpl=gpl.gpl 
        WHERE (gse.title like '%epatit%' OR gse.title like '%HCV%')
            AND gpl.manufacturer like '%Illumina%'
            AND gse.contact like '%Yale%';")
res_v2

# also we can use pandoc table to present this table in a different way
pandoc.table(res[,1:5],style='grid')
```


## Problem 2

```{r}
# reproduce the query using data.table
res.2 <- data.table(dbGetQuery(geo_con, "SELECT 
            gse.title, 
            gse.gse, 
            gpl.gpl, 
            gpl.manufacturer, 
            gpl.title 
        FROM (gse JOIN gse_gpl ON gse.gse=gse_gpl.gse) j 
        JOIN gpl ON j.gpl=gpl.gpl 
        WHERE gse.title like '%HCV%'
            AND gpl.manufacturer like '%Illumina%' 
            AND gse.contact like '%Yale%';"))
res.2
```

We can also use `data.table` to do the joins rather than our SQL query. It might look something like this:

```{r}
# reproduce query using data.table joins instead of SQL joins
res2 <- data.table(dbGetQuery(geo_con, "SELECT gpl, manufacturer, title 
                              from gpl 
                              WHERE manufacturer like '%Illumina%';"), 
                          key="gpl")[setkey(data.table(dbGetQuery(geo_con, 
                              "SELECT * from gse_gpl;"), 
                          key=c("gse", "gpl"))[data.table(dbGetQuery(geo_con, 
                              "SELECT title, gse 
                              from gse 
                              WHERE contact like '%Yale%' 
                              AND title like '%HCV%';"), 
                          key="gse"),nomatch=0],gpl),nomatch=0]

res2

# and it looks the same as our original result:
res
res2
```

## Cleanup
```{r}
# close connnection
dbDisconnect(geo_con)
```
